LLM Output Compliance with Handcrafted Linguistic Features --- An Experiment

### 2025-01-15 article/main.tex
prepare ICAART camera ready submission

\usepackage{algorithm2e}
\usepackage{array}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsthm}
\usepackage{apalike}
\usepackage{balance}
\usepackage{booktabs}
\usepackage{calc}
\usepackage[justification=centering]{caption}
\usepackage{cleveref}
\usepackage{comment}
\usepackage{enumitem}
\usepackage{epsfig}
\usepackage{fancyhdr}
\usepackage[T1]{fontenc}
\usepackage[bottom]{footmisc}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{inconsolata}
\usepackage[utf8]{inputenc}
\usepackage{interval}
\usepackage{microtype}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{pslatex}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{tikz}
\usepackage{url}
\usepackage{verbatim}
\usepackage{xcolor}
\usepackage{xstring}

\usepackage[true]{anonymous-acm}

@@

\begin{document}

\title{LLM Output Compliance with Handcrafted Linguistic Features --- An Experiment}

\author{\authorname{Andrei Olar\sup{1}\orcidAuthor{0009-0006-7913-9276}}
\affiliation{\sup{1}Mathematics and Computer Science Faculty, Babeș-Bolyai University, Cluj-Napoca, România}
\email{andrei.olar@ubbcluj.ro}
}
\keywords{handcrafted linguistic features, large language models, controlled text generation, text style transfer}
\abstract{
    Can we control the writing style of large language models (LLMs) by specifying
    desired linguistic features?
    We address this question by investigating the impact of handcrafted linguistic
    feature (HLF) instructions on LLM-generated text.
    Our experiment evaluates various state-of-the-art LLMs using prompts
    incorporating HLF statistics derived from corpora of CNN articles and Yelp
    reviews.
    We find that LLMs demonstrate sensitivity to these instructions, particularly
    when tasked with conforming to concrete features like word count.
    However, compliance with abstract features, such as lexical variation, proves
    more challenging, often resulting in negative impacts on compliance.
    Our findings highlight the potential and limitations of utilizing HLFs for
    guiding LLM text generation and underscore the need for further research into
    optimizing prompt design and feature selection.
}

\onecolumn \maketitle \normalsize \setcounter{footnote}{0} \vfill

@@

If syntax differs from meaning~\cite{chomsky2002syntactic} the intuition is that
writing style is defined in terms of both.
The distinction is important in order to set boundaries.
HLFs are largely focused on syntax and similarly orderly constructs.
Therefore, we cover the influence of meaning on author style minimally and
superficially at best.

@@

Our main focus is to understand how susceptible LLMs are to prompts containing
instructions to conform to HLFs.
In this experiment we use LFTK~\cite{lftk-2023}, a framework built on top of
spaCy~\cite{spacy} which provides implementations for computing many HLFs.
LFTK categorizes HLFs by domain and family.
We select HLFs based on their domain and family from the catalog implemented by
LFTK, with the interest to make a diverse selection with respect to both
these attributes.
Besides the domain and family, we also choose features based on their influence
over text style as recognized in literature~\cite{verma2019lexical,lugea2023stylistics}.

@@

\begin{table*}[ht!]
    \setlength\tabcolsep{6pt}
    \centering
    \resizebox{\textwidth}{!}{%
        \begin{tabular}{@{}llllll@{}}
            \toprule
            LFTK ID         & Name                              & Family           & Domain           & Abstraction     & Style Level\footnotemark\\ \toprule
            t\_word         & total words                       & wordsent         & surface          & very concrete   & lexical                 \\
            t\_sent         & total sentences                   & wordsent         & surface          & very concrete   & syntactic               \\ \bottomrule
            n\_uverb        & total number of unique verbs      & partofspeech     & syntax           & concrete        & syntactic               \\
            n\_uadj         & total number of unique adjectives & partofspeech     & syntax           & concrete        & syntactic               \\ \bottomrule
            simp\_ttr       & simple type token ratio           & typetokenratio   & lexico-semantics & regular         & pragmatic               \\
            a\_verb\_pw     & number of verbs per word          & partofspeech     & surface          & regular         & syntactic               \\ \bottomrule
            corr\_adj\_var  & corrected adjective variation     & lexicalvariation & lexico-semantics & abstract        & lexical                 \\
            corr\_verb\_var & corrected verb variation          & lexicalvariation & lexico-semantics & abstract        & lexical                 \\ \bottomrule
            fkgl            & Flesch-Kincaid grade level        & readformula      & surface          & highly abstract & pragmatic               \\
            a\_kup\_pw      & Kuperman age of acquisition       & worddiff         & lexico-semantics & highly abstract & lexical                 \\ \bottomrule
        \end{tabular}%
    }
    \caption{Handcrafted Linguistic Features}\label{table-hlf}
\end{table*}
\footnotetext{as described in the figure from~\cite{lugea2023stylistics}, chapter 1, page 7}

@@

The Huggingface ecosystem~\cite{lhoest-etal-2021-datasets} makes the selected
datasets easily accessible.
Yelp reviews are loaded from \texttt{Yelp/yelp\_review\_full}, while CNN stories
are loaded from \texttt{abisee/cnn\_dailymail}.

@@

The result tables show $Diff_N$ for each LLM in the context of a prompt and a
text corpus.
While the main body of the paper is focused on the CNN dataset and, an appendix
to the paper documents the results collected by running the experiment for each
dataset and prompt variant.

@@

\begin{figure*}[ht!]
    \centering
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-claude3-cnn_dailymail/prompt_2-claude3-cnn_dailymail_n_uadj.png}
        \caption{Claude3\texttt{n\_uadj}\\CNN-DailyMail, Scenario 2, Input Inside Corpus}\label{fig-p2-ifd-claude3-nuadj}
    \end{minipage}
    \hfill
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail_t_word.png}
        \caption{Gemini\texttt{t\_word}\\CNN-DailyMail, Scenario 2, Input Inside Corpus}\label{fig-p2-ifd-gemini-twords}
    \end{minipage}
    \hfill
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-llama3_70b-cnn_dailymail/prompt_2-llama3_70b-cnn_dailymail_a_kup_pw.png}
        \caption[center]{Llama3\texttt{a\_kup\_pw}\\CNN-DailyMail, Scenario 2, Input Inside Corpus}\label{fig-p2-ifd-llama3-a-kup-pw}
    \end{minipage}
\end{figure*}

@@

\bibliography{references}

\appendix

\section{Additional Result Tables}\label{sec:yelp-tables}
\subsection{Yelp}
\Cref{table-prompt-1-yelp,table-prompt-2-yelp,table-prompt-2-ifd-yelp} show the
experiment results on the corpus of reviews extracted from the Yelp dataset.

\begin{table*}[ht]
    \centering
    \resizebox{\textwidth}{!}{%
        \begin{tabular}{lllllllllll}
            model                 & t\_word                   & t\_sent                    & n\_uverb                 & n\_uadj                  & simp\_ttr & a\_verb\_pw             & corr\_adj\_var          & corr\_verb\_var         & fkgl                       & a\_kup\_pw               \\
            \toprule
            claude3 $Diff_N$     & \textbf{\textit{-281.9}}  & \textbf{\textit{-13.22}}   & \textbf{\textit{-25.79}} & -14.56                   & -0.06     & 0.0                     & -1.27                   & -1.14                   & \textbf{\underline{6.89}}  & \textbf{\underline{1.0}} \\ \midrule
            claude3 $Diff_A$     & \textbf{\textit{-774.0}}  & \textbf{\textit{-27.76}}   & \textbf{\textit{-70.0}}  & -38.0                    & -0.19     & 0.02                    & -3.25                   & -3.39                   & \textbf{\underline{21.68}} & \textbf{\underline{2.9}} \\ \midrule
            gemini $Diff_N$      & -109.27                   & 0.05                       & -7.25                    & -13.18                   & -0.14     & 0.01                    & -1.01                   & -0.4                    & 1.62                       & -0.4                     \\ \midrule
            gemini $Diff_A$      & -252.82                   & -0.78                      & -17.5                    & -23.14                   & -0.52     & 0.01                    & -1.85                   & -1.07                   & 7.14                       & -0.93                    \\ \midrule
            gpt $Diff_N$         & -21.62                    & 2.42                       & -3.8                     & -8.84                    & 0.0       & -0.01                   & -0.8                    & -0.53                   & 8.94                       & -0.37                    \\ \midrule
            gpt $Diff_A$         & -34.72                    & 9.68                       & -10.54                   & -24.57                   & -0.03     & -0.02                   & -2.37                   & -1.58                   & 19.26                      & -1.03                    \\ \midrule
            command\_r $Diff_N$  & -40.1                     & 1.22                       & 2.43                     & \textbf{\textit{-13.64}} & 0.05      & \textbf{\textit{-0.04}} & \textbf{\textit{-1.17}} & 0.59                    & \textbf{\textit{-7.25}}    & \textbf{\textit{-1.67}}  \\ \midrule
            command\_r $Diff_A$  & -75.5                     & 1.5                        & 12.94                    & \textbf{\textit{-34.0}}  & 0.19      & \textbf{\textit{-0.12}} & \textbf{\textit{-2.76}} & 2.83                    & \textbf{\textit{-20.49}}   & \textbf{\textit{-4.76}}  \\ \midrule
            llama3\_70b $Diff_N$ & \textbf{\textit{-210.45}} & \textbf{\underline{6.87}}  & \textbf{\textit{-13.46}} & \textbf{\textit{-15.17}} & -0.16     & \textbf{\textit{-0.02}} & \textbf{\textit{-1.78}} & \textbf{\textit{-1.31}} & 3.7                        & \textbf{\textit{-1.58}}  \\ \midrule
            llama3\_70b $Diff_A$ & \textbf{\textit{-607.52}} & \textbf{\underline{19.42}} & \textbf{\textit{-41.22}} & \textbf{\textit{-44.77}} & -0.42     & \textbf{\textit{-0.05}} & \textbf{\textit{-5.48}} & \textbf{\textit{-4.34}} & 9.79                       & \textbf{\textit{-4.98}}  \\ \bottomrule
        \end{tabular}%
    }
    \caption{Yelp Results - Prompt Without Examples}\label{table-prompt-1-yelp} % chktex-file 8
\end{table*}

\begin{table*}[!ht]
    \centering
    \resizebox{\textwidth}{!}{%
        \begin{tabular}{lllllllllll}
            model                 & t\_word & t\_sent & n\_uverb & n\_uadj & simp\_ttr & a\_verb\_pw & corr\_adj\_var            & corr\_verb\_var & fkgl                       & a\_kup\_pw \\
            \toprule
            claude3 $Diff_N$     & -121.77 & -4.61   & -9.9     & 18.41   & -0.11     & 0.03        & \textbf{\underline{1.4}}  & -0.33           & 1.81                       & 0.95       \\ \midrule
            claude3 $Diff_A$     & -394.5  & -11.5   & -37.5    & 60.0    & -0.37     & 0.09        & \textbf{\underline{4.66}} & -1.49           & 5.38                       & 2.62       \\ \midrule
            gemini $Diff_N$      & -23.49  & 1.52    & -0.55    & -1.29   & -0.02     & -0.01       & 0.09                      & -0.05           & \textbf{\underline{6.41}}  & 0.2        \\ \midrule
            gemini $Diff_A$      & -83.0   & 3.0     & -2.5     & -6.36   & -0.0      & -0.03       & 0.26                      & -0.08           & \textbf{\underline{19.63}} & 0.53       \\ \midrule
            gpt $Diff_N$         & 13.66   & 2.83    & -3.65    & -3.19   & 0.04      & -0.04       & -0.24                     & -0.22           & 7.5                        & 0.0        \\ \midrule
            gpt $Diff_A$         & 55.5    & 9.5     & -9.86    & -12.86  & 0.14      & -0.11       & -0.84                     & -0.22           & 18.67                      & -0.38      \\ \midrule
            command\_r $Diff_N$  & -188.71 & 0.69    & -10.73   & 1.17    & -0.23     & -0.01       & \textbf{\underline{0.74}} & -0.83           & \textbf{\textit{-5.02}}    & -0.08      \\ \midrule
            command\_r $Diff_A$  & -516.82 & 2.02    & -29.22   & 8.28    & -0.66     & -0.0        & \textbf{\underline{2.98}} & -2.48           & \textbf{\textit{-14.52}}   & -0.48      \\ \midrule
            llama3\_70b $Diff_N$ & -202.95 & -7.84   & -13.73   & 1.43    & -0.15     & -0.04       & 0.38                      & -1.39           & 10.67                      & 1.37       \\ \midrule
            llama3\_70b $Diff_A$ & -561.24 & -18.78  & -49.94   & 5.07    & -0.38     & -0.14       & 1.62                      & -4.81           & 29.26                      & 3.11       \\ \bottomrule
        \end{tabular}%
    }
    \caption{Yelp Results - Prompt With Examples\\Input External To Corpus}\label{table-prompt-2-yelp}
\end{table*}
\begin{table*}[!ht]
    \centering
    \resizebox{\textwidth}{!}{%
        \begin{tabular}{lllllllllll}
            model                 & t\_word                   & t\_sent                 & n\_uverb                  & n\_uadj                  & simp\_ttr               & a\_verb\_pw & corr\_adj\_var          & corr\_verb\_var         & fkgl                    & a\_kup\_pw              \\
            \toprule
            claude3 $Diff_N$     & -148.13                   & -1.63                   & -3.93                     & \textbf{\textit{-46.03}} & -0.11                   & -0.0        & \textbf{\textit{-3.08}} & -0.31                   & -4.3                    & \textbf{\textit{-1.53}} \\ \midrule
            claude3 $Diff_A$     & -526.14                   & -7.98                   & -23.86                    & \textbf{\textit{-133.5}} & -0.41                   & -0.0        & \textbf{\textit{-8.98}} & -1.75                   & -21.03                  & \textbf{\textit{-5.34}} \\ \midrule
            gemini $Diff_N$      & 74.27                     & 0.65                    & 4.67                      & 2.98                     & 0.04                    & -0.0        & -0.22                   & 1.05                    & 1.44                    & -0.65                   \\ \midrule
            gemini $Diff_A$      & 221.0                     & 2.0                     & 10.5                      & 7.44                     & 0.12                    & -0.03       & -0.85                   & 2.73                    & 3.28                    & -1.79                   \\ \midrule
            gpt $Diff_N$         & 0.65                      & 4.03                    & -1.19                     & -2.66                    & -0.03                   & -0.01       & -0.17                   & -0.13                   & 1.05                    & -0.67                   \\ \midrule
            gpt $Diff_A$         & 17.0                      & 11.68                   & -3.68                     & -7.0                     & -0.1                    & -0.04       & -0.43                   & -0.05                   & 0.57                    & -1.75                   \\ \midrule
            command\_r $Diff_N$  & \textbf{\textit{-236.11}} & \textbf{\textit{-3.6}}  & \textbf{\textit{-0.97}}   & \textbf{\textit{-27.04}} & \textbf{\textit{-0.23}} & 0.0         & -1.24                   & \textbf{\textit{-1.08}} & \textbf{\textit{-2.13}} & \textbf{\textit{-1.09}} \\ \midrule
            command\_r $Diff_A$  & \textbf{\textit{-727.0}}  & \textbf{\textit{-9.24}} & \textbf{\underline{0.76}} & \textbf{\textit{-68.5}}  & \textbf{\textit{-0.64}} & 0.02        & -2.89                   & \textbf{\textit{-3.08}} & \textbf{\textit{-5.5}}  & \textbf{\textit{-3.0}}  \\ \midrule
            llama3\_70b $Diff_N$ & \textbf{\textit{-333.03}} & \textbf{\textit{-5.19}} & 4.43                      & \textbf{\textit{-21.38}} & -0.16                   & -0.03       & \textbf{\textit{-1.77}} & -0.03                   & 1.87                    & -0.44                   \\ \midrule
            llama3\_70b $Diff_A$ & \textbf{\textit{-687.34}} & \textbf{\textit{-3.42}} & 10.7                      & \textbf{\textit{-61.64}} & -0.31                   & -0.04       & \textbf{\textit{-5.6}}  & -0.89                   & 3.0                     & -1.92                   \\ \bottomrule
        \end{tabular}%
    }
    \caption{Yelp Results - Prompt With Examples\\Input From Corpus}\label{table-prompt-2-ifd-yelp}
\end{table*}

\section{Additional Plots}

\subsection{GPT}

\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-gpt-cnn_dailymail/prompt_1-gpt-cnn_dailymail.png}
    \caption{GPT on CNN Corpus\\Prompt Without Examples}\label{fig:gpt-prompt1-cnn}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-gpt-yelp/prompt_1-gpt-yelp.png}
    \caption{GPT on Yelp Corpus\\Prompt Without Examples}\label{fig:gpt-prompt1-yelp}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-gpt-cnn_dailymail/prompt_2-gpt-cnn_dailymail.png}
    \caption{GPT on CNN Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:gpt-prompt2-cnn}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-gpt-yelp/prompt_2-gpt-yelp.png}
    \caption{GPT on Yelp Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:gpt-prompt2-yelp}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-gpt-cnn_dailymail/prompt_2-gpt-cnn_dailymail.png}
    \caption{GPT on CNN Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:gpt-prompt2-cnn-ifd}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-gpt-yelp/prompt_2-gpt-yelp.png}
    \caption{GPT on Yelp Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:gpt-prompt2-yelp-ifd}
\end{figure*}

\Cref{fig:gpt-prompt1-cnn,fig:gpt-prompt1-yelp,fig:gpt-prompt2-cnn,fig:gpt-prompt2-yelp,fig:gpt-prompt2-cnn-ifd,fig:gpt-prompt2-yelp-ifd}
show the experiment results for the Claude 3 Opus large language model.

\subsection{Gemini}

\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-gemini-cnn_dailymail/prompt_1-gemini-cnn_dailymail.png}
    \caption{Gemini on CNN Corpus\\Prompt Without Examples}\label{fig:gemini-prompt1-cnn}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-gemini-yelp/prompt_1-gemini-yelp.png}
    \caption{Gemini on Yelp Corpus\\Prompt Without Examples}\label{fig:gemini-prompt1-yelp}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail.png}
    \caption{Gemini on CNN Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:gemini-prompt2-cnn}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-gemini-yelp/prompt_2-gemini-yelp.png}
    \caption{Gemini on Yelp Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:gemini-prompt2-yelp}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail.png}
    \caption{Gemini on CNN Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:gemini-prompt2-cnn-ifd}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-gemini-yelp/prompt_2-gemini-yelp.png}
    \caption{Gemini on Yelp Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:gemini-prompt2-yelp-ifd}
\end{figure*}

\Cref{fig:gemini-prompt1-cnn,fig:gemini-prompt1-yelp,fig:gemini-prompt2-cnn,fig:gemini-prompt2-yelp,fig:gemini-prompt2-cnn-ifd,fig:gemini-prompt2-yelp-ifd}
show the experiment results for the Claude 3 Opus large language model.

\subsection{Claude3}

\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-claude3-cnn_dailymail/prompt_1-claude3-cnn_dailymail.png}
    \caption{Claude3 on CNN Corpus\\Prompt Without Examples}\label{fig:claude3-prompt1-cnn}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-claude3-yelp/prompt_1-claude3-yelp.png}
    \caption{Claude3 on Yelp Corpus\\Prompt Without Examples}\label{fig:claude3-prompt1-yelp}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-claude3-cnn_dailymail/prompt_2-claude3-cnn_dailymail.png}
    \caption{Claude3 on CNN Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:claude3-prompt2-cnn}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-claude3-yelp/prompt_2-claude3-yelp.png}
    \caption{Claude3 on Yelp Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:claude3-prompt2-yelp}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-claude3-cnn_dailymail/prompt_2-claude3-cnn_dailymail.png}
    \caption{Claude3 on CNN Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:claude3-prompt2-cnn-ifd}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-claude3-yelp/prompt_2-claude3-yelp.png}
    \caption{Claude3 on Yelp Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:claude3-prompt2-yelp-ifd}
\end{figure*}

\Cref{fig:claude3-prompt1-cnn,fig:claude3-prompt1-yelp,fig:claude3-prompt2-cnn,fig:claude3-prompt2-yelp,fig:claude3-prompt2-cnn-ifd,fig:claude3-prompt2-yelp-ifd}
show the experiment results for the Claude 3 Opus large language model.

\subsection{Llama3}

\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-llama3_70b-cnn_dailymail/prompt_1-llama3_70b-cnn_dailymail.png}
    \caption{Llama3 on CNN Corpus\\Prompt Without Examples}\label{fig:llama3_70b-prompt1-cnn}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_1/prompt_1-llama3_70b-yelp/prompt_1-llama3_70b-yelp.png}
    \caption{Llama3 on Yelp Corpus\\Prompt Without Examples}\label{fig:llama3_70b-prompt1-yelp}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-llama3_70b-cnn_dailymail/prompt_2-llama3_70b-cnn_dailymail.png}
    \caption{Llama3 on CNN Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:llama3_70b-prompt2-cnn}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2/prompt_2-llama3_70b-yelp/prompt_2-llama3_70b-yelp.png}
    \caption{Llama3 on Yelp Corpus\\Prompt With Examples\\Extraneous Input}\label{fig:llama3_70b-prompt2-yelp}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-llama3_70b-cnn_dailymail/prompt_2-llama3_70b-cnn_dailymail.png}
    \caption{Llama3 on CNN Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:llama3_70b-prompt2-cnn-ifd}
\end{figure*}
\begin{figure*}[ht]
    \includegraphics[width=\textwidth,height=0.9\textheight,scale=1]{plots/prompt_2_ifd/prompt_2-llama3_70b-yelp/prompt_2-llama3_70b-yelp.png}
    \caption{Llama3 on Yelp Corpus\\Prompt With Examples\\Input from Corpus}\label{fig:llama3_70b-prompt2-yelp-ifd}
\end{figure*}

\Cref{fig:llama3_70b-prompt1-cnn,fig:llama3_70b-prompt1-yelp,fig:llama3_70b-prompt2-cnn,fig:llama3_70b-prompt2-yelp,fig:llama3_70b-prompt2-cnn-ifd,fig:llama3_70b-prompt2-yelp-ifd}
show the experiment results for the Claude 3 Opus large language model.

\subsection{Command R+}

>>>>

\usepackage{algorithm2e}
\usepackage{array}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsthm}
\usepackage{apalike}
\usepackage{balance}
\usepackage{booktabs}
\usepackage{calc}
\usepackage[justification=centering]{caption}
\usepackage{comment}
\usepackage{enumitem}
\usepackage{epsfig}
\usepackage{fancyhdr}
\usepackage[T1]{fontenc}
\usepackage[bottom]{footmisc}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{inconsolata}
\usepackage[utf8]{inputenc}
\usepackage{interval}
\usepackage{microtype}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{pslatex}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{tikz}
\usepackage{url}
\usepackage{verbatim}
\usepackage{xcolor}
\usepackage{xstring}

\usepackage[false]{anonymous-acm}

@@

\begin{document}
\title{LLM Output Compliance with Handcrafted Linguistic Features --- An Experiment}
\author{
    \authorname{Andrei Olar\sup{1}\orcidAuthor{0009-0006-7913-9276}}
    \affiliation{\sup{1}Mathematics and Computer Science Faculty, Babeș-Bolyai University, Cluj-Napoca, România}
    \email{andrei.olar@ubbcluj.ro}
}
\keywords{handcrafted linguistic features, large language models, controlled text generation, text style transfer}
\abstract{
    Can we control the writing style of large language models (LLMs) by specifying
    desired linguistic features?
    We address this question by investigating the impact of handcrafted linguistic
    feature (HLF) instructions on LLM-generated text.
    Our experiment evaluates various state-of-the-art LLMs using prompts
    incorporating HLF statistics derived from corpora of CNN articles and Yelp
    reviews.
    We find that LLMs demonstrate sensitivity to these instructions, particularly
    when tasked with conforming to concrete features like word count.
    However, compliance with abstract features, such as lexical variation, proves
    more challenging, often resulting in negative impacts on compliance.
    Our findings highlight the potential and limitations of utilizing HLFs for
    guiding LLM text generation and underscore the need for further research into
    optimizing prompt design and feature selection.
}
\onecolumn \maketitle \normalsize \setcounter{footnote}{0} \vfill
\interfootnotelinepenalty=10000

@@

Syntax differs from meaning~\cite{chomsky2002syntactic} and the intuition is
that writing style is defined in terms of both.
The distinction is important in order to set boundaries.
HLFs are largely focused on syntax and similarly orderly constructs.
Therefore, we cover the influence of meaning on author style minimally and
superficially at best.

@@

Our main focus is to understand how susceptible LLMs are to prompts containing
instructions to conform to HLFs.
In this experiment we use LFTK~\cite{lftk-2023}, a framework built on top of
spaCy~\cite{spacy} which provides implementations for computing many HLFs.
LFTK categorizes HLFs by domain and family.
We select HLFs based on their domain and family from the catalog implemented by
LFTK, with the interest to make a diverse selection with respect to both
these attributes.
Besides the domain and family, we also choose features based on the
level\footnotemark~at which they influence text style as recognized in
literature~\cite{verma2019lexical,lugea2023stylistics}.
\footnotetext{the notion of style level is described
in~\cite{lugea2023stylistics}, chapter 1, page 7}

@@

\begin{table*}[h!]
    \setlength\tabcolsep{6pt}
    \centering
    \resizebox{\textwidth}{!}{%
        \begin{tabular}{@{}llllll@{}}
            \toprule
            LFTK ID         & Name                              & Family           & Domain           & Abstraction     & Style Level \\ \toprule
            t\_word         & total words                       & wordsent         & surface          & very concrete   & lexical     \\
            t\_sent         & total sentences                   & wordsent         & surface          & very concrete   & syntactic   \\ \bottomrule
            n\_uverb        & total number of unique verbs      & partofspeech     & syntax           & concrete        & syntactic   \\
            n\_uadj         & total number of unique adjectives & partofspeech     & syntax           & concrete        & syntactic   \\ \bottomrule
            simp\_ttr       & simple type token ratio           & typetokenratio   & lexico-semantics & regular         & pragmatic   \\
            a\_verb\_pw     & number of verbs per word          & partofspeech     & surface          & regular         & syntactic   \\ \bottomrule
            corr\_adj\_var  & corrected adjective variation     & lexicalvariation & lexico-semantics & abstract        & lexical     \\
            corr\_verb\_var & corrected verb variation          & lexicalvariation & lexico-semantics & abstract        & lexical     \\ \bottomrule
            fkgl            & Flesch-Kincaid grade level        & readformula      & surface          & highly abstract & pragmatic   \\
            a\_kup\_pw      & Kuperman age of acquisition       & worddiff         & lexico-semantics & highly abstract & lexical     \\ \bottomrule
        \end{tabular}%
    }
    \caption{Handcrafted Linguistic Features}\label{table-hlf}
\end{table*}

@@

The datasets are available in the Huggingface
ecosystem~\cite{lhoest-etal-2021-datasets}.
Yelp reviews are loaded from \texttt{Yelp/yelp\_review\_full}, while CNN stories
are loaded from \texttt{abisee/cnn\_dailymail}.

@@

The result tables show $Diff_N$ for each LLM in the context of a prompt and a
text corpus.
The paper is focused on some key results observed on the CNN article corpus.
Additional results are available
\textlinkanon{https://github.com/koosie0507/llm-tst-consistency/tree/main/reports}{online}.

@@

\begin{figure*}[ht!]
    \centering
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-claude3-cnn_dailymail/prompt_2-claude3-cnn_dailymail_n_uadj.png}
        \caption{Claude3~\texttt{n\_uadj}\\CNN-DailyMail, Scenario 2, Input Inside Corpus}\label{fig-p2-ifd-claude3-nuadj}
    \end{minipage}
    \hfill
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-gemini-cnn_dailymail/prompt_2-gemini-cnn_dailymail_t_word.png}
        \caption{Gemini~\texttt{t\_word}\\CNN-DailyMail, Scenario 2, Input Inside Corpus}\label{fig-p2-ifd-gemini-twords}
    \end{minipage}
    \hfill
    \begin{minipage}{0.32\textwidth}
        \includegraphics[width=\linewidth]{plots/prompt_2_ifd/prompt_2-llama3_70b-cnn_dailymail/prompt_2-llama3_70b-cnn_dailymail_a_kup_pw.png}
        \caption[center]{Llama3~\texttt{a\_kup\_pw}\\CNN-DailyMail, Scenario 2, Input Inside Corpus}\label{fig-p2-ifd-llama3-a-kup-pw}
    \end{minipage}
\end{figure*}

@@

\bibliographystyle{apalike}
{\small
\bibliography{references}}
\balance
\end{document}
